Error Bounds for Extended Source Inversion applied to an Acoustic Transmission Inverse Problem

### 2024-02-15 repro/papers/discreptheorypub/paper.tex
more
cleaned up

\title{Error Bounds for Extended Source Inversion applied to an Acoustic Transmission Inverse Problem}
\author[Symes]{William W. Symes$^1$\\
$^1$ PO Box 43, Orcas, WA 98280 USA, {\tt symes@rice.edu}}

\maketitle
\begin{abstract}
  A simple inverse problem for the wave equation requires
  determination of both the wave velocity in a homogenous acoustic
  material and the transient waveform of an isotropic point radiator,
  given the time history of the wavefield at a remote point in
  space. The duration (support) of the source waveform and the
  source-to-receiver distance are assumed known. A least squares
  formulation of this problem exhibits the ``cycle-skipping''
  behaviour observed in field scale problems of this type, with many
  local minima differing greatly from the global minimizer. An
  extended formulation, dropping the support constraint on the source
  waveform in favor of a weighted quadratic penalty, eliminates this
  misbehaviour. With proper choice of the weight operator, the
  velocity component at {\em any} local minimizer of this extended
  objective function differs from the global
  minimizer of the least-squares formulation by less than a linear
  combination of the source waveform support radius and data
  noise-to-signal ratio.

@@

\section{An Acoustic transmission Inverse Problem}
According to linear acoustics, the causal pressure field $p(\bx,t)$ due to an
isotropic point radiator at $\bx_s \in \bR^3$ with time-varying
intensity (``wavelet'') $w(t)$, propagating in an elastic fluid with
slowness (reciprocal velocity) $m$ and density $1$ (in appropriate
units), is the solution of the the initial value problem for the wave equation \cite[]{Frie:58}:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0.
\end{eqnarray}
The solution is well-known, see for instance
\cite{CourHil:62}, Chapter VI, section 12, equation 47:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right),
\end{equation}
This trace of $p$ at $\bx_r \in \bR^3$ can be viewed as the result of
applying an m-dependent linear operator $F[m]$ to the wavelet $w$:
\begin{equation}
\label{eqn:mod}
F[m]w(t)  = p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right) 
\end{equation}
The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km. The pressure trace
$p(\bx_r,\cdot)$ is square-integrable if and only if the same is true
of the wavelet $w$. Since the square-integral of the pressure trace is
proportional to the accumulated energy transferred from the fluid to
the sensor \cite[]{SantosaSymes:00}, assume that $w \in L^2(\bR)$.

@@

Define the relative mean-square error $\lerr: M \times \lW \times D
\rightarrow \bR^+$ by
\begin{equation}
  \label{eqn:redms}
  \lerr[m,w;d]=\frac{1}{2}\|\lF[m]w-d\|^2/\|d\|^2,
\end{equation}
so that inequality \ref{eqn:probstat0} is equivalent to
\begin{equation}
  \label{eqn:probstat1}
  \lerr[m,w;d] \le \frac{1}{2}\epsilon^2,
\end{equation}
Minimization of $\lerr$ is the standard least-squares formulation of
the Inverse Problem defined above: if the global minimum value of
$\lerr$ is less than $\frac{1}{2}\epsilon^2$, then the global
minimizer is a solution of the Inverse Problem.

@@

\noindent{\bf Remark:} A useful consequence of the condition 
\ref{eqn:fullrec}: for every $m \in M$, 
\begin{equation}
  [-\lambda_{\rm max}, -\lambda_{\rm max}] \subset[ t_{\rm min}-mr , 
  t_{\rm max}-mr]. 
  \label{eqn:zeroinc}
\end{equation}

@@

\begin{proof} From the definition \ref{eqn:mod},
\[
 \lerr[m,w;d] =  \frac{1}{32\pi^2
    r^2\|d\|^2}\int\,dt\,\left|w\left(t-mr\right)-w_*\left(t-m_*r\right)\right|^2
\]
Since $w_*, w$ vanish for $|t|>\lambda$,
$\lF[m_*]w_*(t)$ vanishes if $|t-m_*r|>\lambda$ and $\lF[m]w$ vanishes if $|t-mr|>\lambda$. So if $|mr-m_*r|
= |m-m_*|r > 2\lambda$, then $|t-mr|+|t-m_*r| \ge |mr-m_*r| >
2\lambda$ so either $|t-mr|>\lambda$ or $|t-m_*r|>\lambda$, that is,
either $\lF[m]w(t)=0$ or $\lF[m_*]w_*=0$. Therefore $\lF[m]w$ and
$\lF[m_*]w_*$ are orthogonal in the sense of the $L^2$ inner product
$\langle \cdot,\cdot \rangle_D$ on $D$:
\begin{equation}
  \label{eqn:ortho}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \langle F[m]w,
  F[m_*]w_*\rangle_D = 0
\end{equation}
But $d = \lF[m_*]w_*$, so this is the same as saying that $d$ is
orthogonal to $F[m]w$. So conclude after a minor manipulation that
\[
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \lerr[m,w;d]=\frac{1}{32\pi^2 
    r^2\|d\|^2}(\|w\|^2 + \|w_*\|^2).
\]
\begin{equation}
  \label{eqn:iso}
  = \frac{1}{2}\left(\frac{\|w\|^2}{\|w_*\|^2} + 1 \right)
\end{equation}
That is, for slowness $m$ in error by more than $2\lambda/r$ from the 
target slowness $m_*$, the means square error (FWI objective) $\lerr$ is independent of
$m$, and its minimum over $w$ is attained for $w=0$
\end{proof}

@@

The main theoretical device used in the proofs of our main results on 
extended inversion is Variable Projection reduction of the penalty objective $\Ja$
(equation \ref{eqn:jdef}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$:
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}
A minimizer $w$ on the right-hand side of definition
\ref{eqn:redexp} must solve the {\em normal equation}
\begin{equation}
  \label{eqn:norm}
  (F[m]^TF[m]+\alpha^2A^TA)w= F[m]^Td, 
\end{equation}
(here and in the following, the superscript $T$ signifies adjoint in
the sense of domain and range Hilbert structures).

@@

\begin{proposition}
  \label{thm:norminvexp}
  Assume the conditions \ref{eqn:mod}, \ref{eqn:fullrec},
  \ref{eqn:annmult}. Also assume that $\lambda \in (0,\lambda_{\rm
    max}], \alpha > 0,$ and that  $C>0$ exists so that $a \in L^{\infty}(\bR)$
  mentioned in equation \ref{eqn:annmult} satisfies the condition
\begin{equation}
    \label{eqn:abnd} 
    a \ge 0; \, a(t) \ge C\mbox{ for }|t| \ge \lambda_{\rm max}.
  \end{equation}
%  \begin{equation}
%   \label{eqn:abnd}
%    |t| > \lambda \Rightarrow a(t) \ge C,
%  \end{equation}
%  in which $C>0$ may depend on $\lambda$.
  Then
  \begin{itemize}
  \item[1. ]the normal operator $F[m]^TF[m] + \alpha^2A^TA$ is
    invertible for any $m \in M$, $\alpha > 0$;
  \item[2. ]the solution $\aw[m;d]\in W$ of the normal equation
    \ref{eqn:norm} is given by
    \begin{equation}
      \label{eqn:normsol}
      \aw[m;d](t) = \left\{
        \begin{array}{c}
          \left(\frac{1}{(4\pi r)^2} + \alpha^2
          a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
          min}-mr, t_{\rm max}-mr];\\
          0, \mbox{ else;}
        \end{array}
      \right.
    \end{equation}
  \item[3. ]if in addition $d=F[m_*]w_*, w_* \in \lW$ is noise-free, as in equation
    \ref{eqn:defdatanonoise},
    \begin{equation}
      \label{eqn:solnonoise}
      \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 a(t)^2\right)^{-1}w_*\left(t+(m-m_*)r\right).
    \end{equation}
  \end{itemize}
\end{proposition}

\begin{proof} 
  \begin{itemize}
  \item[1. ]Note that thanks to \ref{eqn:zeroinc}, if $|t|\le
    \lambda \le \lambda_{\rm max}$, then ${\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]}(t) = 1$, whereas if $|t|>\lambda$,
    then $a(t) \ge C$, whence
    \[
      \frac{1}{(4\pi r)^2}  {\bf 1}_{[t_{\rm min}-mr,  
        t_{\rm max}-mr]} + \alpha^2a^2  \ge \min\{(4\pi r)^2,
      \alpha^2\min\{1/(4\pi r)^2,C^2\} \}> 0.
    \]
    Therefore the normal operator is invertible under the stated
    conditions.

@@

\begin{proposition}
  \label{thm:epjgen}
  Assume the hypotheses of Proposition \ref{thm:norminvexp}. Then
  \begin{equation}
  \label{eqn:residnormgen}
  e[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r \alpha a(t-mr))^4(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2
\end{equation}
\begin{equation}
  \label{eqn:anninormgen}
  p[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r a(t-mr))^2(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2
\end{equation}

\begin{equation}
  \label{eqn:expjgen}
\tJa[m;d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}}\,dt\,(4\pi r \alpha a(t-mr))^2(1+(4\pi r \alpha 
a(t-mr))^2)^{-1}d(t)^2. 
\end{equation}
\end{proposition}

@@

  That the VPM objective $\tJa$ is given by \ref{eqn:expjgen} follows from equations \ref{eqn:redexp1}, \ref{eqn:residnormgen}, and
  \ref{eqn:anninormgen}.
\end{proof}

@@

It will be useful to record expressions for the various componenets of
$\tJa$ when the data is noise-free, that is, the context of
Proposition \ref{thm:norminvexp}, item 3:
\begin{corollary}
  \label{thm:epjnonoise}
  Assume the hypotheses of Proposition \ref{thm:norminvexp}, item
  3. Then noting that $\|d\| = \|w_*\|/(4 \pi r)$
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= \frac{\alpha^4}{2\|w_*\|^2}\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m_*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{(4\pi r)^2}{2\|w_*\|^2}\int \,dt\,  
  \frac{a(t-(m-m_*)r)^2}{(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2)^{2}}w_*(t)^2.
\end{equation}
so 
\begin{equation}
\label{eqn:expjnonoise}
\tJa[m;d] = \frac{(4\pi r \alpha)^2}{2\|w_*\|^2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-1}w_*(t)^2. 
\end{equation}
Finally, if $a \in W^{1,\infty}(\bR)$, then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjnonoise}
  \frac{d}{dm}\tJa[m;d] = -\frac{r (4\pi r \alpha)^2}{\|w_*\|^2} \int \,dt \, 
  \frac{\left(a\frac{da}{dt}\right)(t-(m-m_*)r)}{(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{2}}w_*(t)^2. 
\end{equation}
\end{corollary}

@@

\begin{proposition}
  \label{thm:rampgood}
  Suppose that
  \begin{enumerate}
  \item $m_* \in M$;
  \item $0 < \mu \le \lambda$, and $w_* \in W_{\mu}$;
  \item $d_* = F[m_*]w_*$;
  \item $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
    with $\tau$ given by equation \ref{eqn:taudef}; and
  \item $\alpha > 0$.
  \end{enumerate}
  Then for any $m \in M$, 
  \begin{equation}
    | (m - m_*)r| > \lambda  \Rightarrow  \left|\frac{d}{dm}\tJa[m;d_*]\right| >  
    \frac{r(4 \pi r \alpha)^2(\lambda-\mu)}{(1+(4\pi r\alpha)^2 
      (\lambda+\mu)^2)^{2}}  
    \label{eqn:gradbndnonoise}
  \end{equation}
\end{proposition}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d_*] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore, $a(t) = |t|$, $a a'(t) = t$ in the
  support of the integrand on the RHS of equation
  \ref{eqn:dexpjnonoise}, which therefore 
  becomes (after change of integration variable)
  %%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{d}{dm}\tJa[m;d_*] = -\frac{r (4\pi r\alpha)^2}{\|w_*\|^2} \int \,dt \, 
  t(1+(4\pi r)^2 \alpha^2 
  t^2)^{-2}w_*(t+(m-m_*))^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -\frac{r(4 \pi r \alpha)^2}{\|w_*\|^2}\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, t(1+(4\pi r)^2\alpha^2 t^2)^{-2}w_*\left(t+(m-m_*)r\right)^2
  \]
  Suppose that $\mu \le \lambda$ and $w_* \in W_{\mu}$. 
  If $m > m_*+\lambda/r$, then $t+(m-m_*)r \in \mbox{supp }w_*$
  implies $-\mu - \lambda < t < \mu-\lambda<0$, so 
  \[
    t(1+(4\pi r)^2\alpha^2 t^2)^{-2} < (\mu-\lambda)(1+(4\pi r)^2\alpha^2 (\mu+\lambda)^2)^{-2}<0
  \]
  in the support of the integrand in equation
  \ref{eqn:gradfinal}. Arguing similarly for $m<m_*-\lambda/r$, obtain
  a similar inequality, implying the conclusion \ref{eqn:gradbndnonoise}.
\end{proof}

@@

The inequality \ref{eqn:estresidnorm} can be interpreted as a bound 
on $\alpha$, given $\epsilon$ and $\lambda$, for a
stationary point of $\tJa$ to yield a solution of the inverse
problem: one obtains a solution, provided that $\alpha$ is
sufficiently small. On the other hand, it is clear that $\alpha$
cannot be too large if stationary points of $\tJa$ are to yield
solutions: the integrand in \ref{eqn:residnorm} is increasing in
$\alpha$ for every $t$ and $m$, and the multiplier
\[
t \mapsto (4\pi r \alpha(t-(m-m_*)r))^4(1+(4\pi r)^2 \alpha^2 
|t-(m-m_*)r|^2)^{-2}
\]
tends monotonically to $1$ as $\alpha \rightarrow \infty$, uniformly
on the complement of any open interval containing
$t=(m-m_*)r$. Therefore
\begin{equation}
  \label{eqn:elimit}
  \lim_{\alpha \rightarrow \infty}e[m,\aw[m;d];d] =
  \frac{1}{2}\frac{1}{(4 \pi r)^2}\|w_*\|^2 = \frac{1}{2}\|d\|^2.
\end{equation}
Consequently, there exists $\alpha_{\rm max}(\epsilon,\lambda,d)$ so
that
\[
  e[m,\aw[m;d];d]  \le \frac{1}{2}\epsilon^2\|d\|^2
  \Rightarrow \alpha \le \alpha_{\rm max}(\epsilon,\lambda,d).
\]
The existence of this limiting penalty weight has been inferred
indirectly. \cite{FuSymes2017discrepancy} describe a constructive
algorithm for its approximation, which is used in the companion paper
\cite{SymesChenMinkoff:21} to dynamically adjust $\alpha$ in the
course of optimization of $\tJa$.

>>>>

\title{Error Bounds for Extended Source Inversion applied to an Acoustic Transmission Inverse Problem}
\author{William W. Symes\\
  Department of Computational and Applied Mathematics, \\
  Rice University, Houston, Texas 77251, USA; {\tt symes@rice.edu}}

\lefthead{Error Bounds for Extended Source Inversion}
\righthead{Symes}

\begin{abstract}
  A simple inverse problem for the wave equation requires
  determination of both the wave velocity in a homogenous acoustic
  material and the transient waveform of an isotropic point radiator,
  given the time history of the wavefield at a remote point in
  space. The duration (support) of the source waveform and the
  source-to-receiver distance are assumed known. A least squares
  formulation of this problem exhibits the ``cycle-skipping''
  behaviour observed in field scale problems of this type, with many
  local minima differing greatly from the global minimizer. An
  extended formulation, dropping the support constraint on the source
  waveform in favor of a weighted quadratic penalty, eliminates this
  misbehaviour. With proper choice of the weight operator, the
  velocity component at {\em any} local minimizer of this extended
  objective function differs from the global
  minimizer of the least-squares formulation by less than a linear
  combination of the source waveform support radius and data
  noise-to-signal ratio.

@@

\section{An Acoustic transmission Inverse Problem}
According to linear acoustics, the causal pressure field $p(\bx,t)$ due to an
isotropic point radiator at $\bx_s \in \bR^3$ with time-varying
intensity (``wavelet'') $w(t)$, propagating in an elastic fluid with
slowness (reciprocal velocity) $m$ and density $1$ (in appropriate
units), is the solution of the the initial value problem for the wave equation \cite[]{Frie:58}:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s), \nonumber\\
  p(\bx,t)&=&0, t\ll 0.
\end{eqnarray}
The solution is well-known, see for instance
\cite{CourHil:62}, Chapter VI, section 12, equation 47:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right).
\end{equation}
This trace of $p$ at $\bx_r \in \bR^3$ can be viewed as the result of
applying an m-dependent linear operator $F[m]$ to the wavelet $w$:
\begin{equation}
\label{eqn:mod}
F[m]w(t)  = p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right).
\end{equation}
The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km. The pressure trace
$p(\bx_r,\cdot)$ is square-integrable if and only if the same is true
of the wavelet $w$. Since the square-integral of the pressure trace is
proportional to the accumulated energy transferred from the fluid to
the sensor \cite[]{SantosaSymes:00}, assume that $w \in L^2(\bR)$.

@@

Define the relative mean-square error $\lerr: M \times \lW \times D
\rightarrow \bR^+$ by
\begin{equation}
  \label{eqn:redms}
  \lerr[m,w;d]=\frac{1}{2}\|\lF[m]w-d\|^2/\|d\|^2,
\end{equation}
so that inequality \ref{eqn:probstat0} is equivalent to
\begin{equation}
  \label{eqn:probstat1}
  \lerr[m,w;d] \le \frac{1}{2}\epsilon^2.
\end{equation}
Minimization of $\lerr$ is the standard least-squares formulation of
the Inverse Problem defined above: if the global minimum value of
$\lerr$ is less than $\frac{1}{2}\epsilon^2$, then the global
minimizer is a solution of the Inverse Problem.

@@

\noindent{\bf Remark:} A useful consequence of the condition 
\ref{eqn:fullrec}: for every $m \in M$, 
\begin{equation}
  [-\lambda_{\rm max}, \lambda_{\rm max}] \subset[ t_{\rm min}-mr , 
  t_{\rm max}+mr]. 
  \label{eqn:zeroinc}
\end{equation}

@@

\begin{proof} From the definition \ref{eqn:mod},
\[
 \lerr[m,w;d] =  \frac{1}{32\pi^2
    r^2\|d\|^2}\int\,dt\,\left|w\left(t-mr\right)-w_*\left(t-m_*r\right)\right|^2.
\]
Since $w_*, w$ vanish for $|t|>\lambda$,
$\lF[m_*]w_*(t)$ vanishes if $|t-m_*r|>\lambda$ and $\lF[m]w$ vanishes if $|t-mr|>\lambda$. So if $|mr-m_*r|
= |m-m_*|r > 2\lambda$, then $|t-mr|+|t-m_*r| \ge |mr-m_*r| >
2\lambda$ so either $|t-mr|>\lambda$ or $|t-m_*r|>\lambda$, that is,
either $\lF[m]w(t)=0$ or $\lF[m_*]w_*=0$. Therefore $\lF[m]w$ and
$\lF[m_*]w_*$ are orthogonal in the sense of the $L^2$ inner product
$\langle \cdot,\cdot \rangle_D$ on $D$:
\begin{equation}
  \label{eqn:ortho}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \langle F[m]w,
  F[m_*]w_*\rangle_D = 0.
\end{equation}
But $d = \lF[m_*]w_*$, so this is the same as saying that $d$ is
orthogonal to $F[m]w$. So conclude after a minor manipulation that
\[
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \lerr[m,w;d]=\frac{1}{32\pi^2 
    r^2\|d\|^2}(\|w\|^2 + \|w_*\|^2)
\]
\begin{equation}
  \label{eqn:iso}
  = \frac{1}{2}\left(\frac{\|w\|^2}{\|w_*\|^2} + 1 \right).
\end{equation}
That is, for slowness $m$ in error by more than $2\lambda/r$ from the 
target slowness $m_*$, the means square error (FWI objective) $\lerr$ is independent of
$m$, and its minimum over $w$ is attained for $w=0$.
\end{proof}

@@

The main theoretical device used in the proofs of our main results on 
extended inversion is reduction of the penalty objective $\Ja$
(equation \ref{eqn:jdef}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$:
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d].
\end{equation}
Reduction of this type is often termed ``Variable Projection'', after \cite{GolubPereyra:73,GolubPereyra:03}.

A minimizer $w$ on the right-hand side of definition
\ref{eqn:redexp} must solve the {\em normal equation}
\begin{equation}
  \label{eqn:norm}
  (F[m]^TF[m]+\alpha^2A^TA)w= F[m]^Td, 
\end{equation}
(here and in the following, the superscript $T$ signifies adjoint in
the sense of domain and range Hilbert structures).

@@

\begin{proposition}
  \label{thm:norminvexp}
  Assume the conditions \ref{eqn:mod}, \ref{eqn:fullrec},
  \ref{eqn:annmult}. Also assume that $\lambda \in (0,\lambda_{\rm
    max}], \alpha > 0,$ and that  $C>0$ exists so that $a \in L^{\infty}(\bR)$
  mentioned in equation \ref{eqn:annmult} satisfies the condition
  \begin{equation}
    \label{eqn:abnd} 
    a \ge 0; \, a(t) \ge C\mbox{ for }|t| \ge \lambda_{\rm max}.
  \end{equation}
%  \begin{equation}
%   \label{eqn:abnd}
%    |t| > \lambda \Rightarrow a(t) \ge C,
%  \end{equation}
%  in which $C>0$ may depend on $\lambda$.
  Then
  \begin{itemize}
  \item[1. ]the normal operator $F[m]^TF[m] + \alpha^2A^TA$ is
    invertible for any $m \in M$, $\alpha > 0$;
  \item[2. ]the solution $\aw[m;d]\in W$ of the normal equation
    \ref{eqn:norm} is given by
    \begin{equation}
      \label{eqn:normsol}
      \aw[m;d](t) = \left\{
        \begin{array}{c}
          \left(\frac{1}{(4\pi r)^2} + \alpha^2
          a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
          min}-mr, t_{\rm max}-mr];\\
          0, \mbox{ else.}
        \end{array}
      \right.
    \end{equation}
  \item[3. ]if in addition $d=F[m_*]w_*, w_* \in \lW$ is noise-free, as in equation
    \ref{eqn:defdatanonoise},
    \begin{equation}
      \label{eqn:solnonoise}
      \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 a(t)^2\right)^{-1}w_*\left(t+(m-m_*)r\right).
    \end{equation}
  \end{itemize}
\end{proposition}

\begin{proof}
  The idea here is that $F$ fails to be coercive just where $A$ is
  coercive, and vis-versa.
  \begin{itemize}
  \item[1. ]Note that thanks to \ref{eqn:zeroinc}, if $|t|\le
    \lambda \le \lambda_{\rm max}$, then ${\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]}(t) = 1$, whereas if $|t|>\lambda$,
    then $a(t) \ge C$, whence
    \[
      \frac{1}{(4\pi r)^2}  {\bf 1}_{[t_{\rm min}-mr,  
        t_{\rm max}-mr]} + \alpha^2a^2  \ge \min\{(4\pi r)^2,
      \alpha^2\min\{1/(4\pi r)^2,C^2\} \}> 0.
    \]
    Therefore the normal operator is invertible under the stated
    conditions.

@@

\begin{proposition}
  \label{thm:epjgen}
  Assume the hypotheses of Proposition \ref{thm:norminvexp}. Then
  \begin{equation}
  \label{eqn:residnormgen}
  e[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r \alpha a(t-mr))^4(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2,
\end{equation}
\begin{equation}
  \label{eqn:anninormgen}
  p[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r a(t-mr))^2(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2,
\end{equation}
and
\begin{equation}
  \label{eqn:expjgen}
\tJa[m;d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}}\,dt\,(4\pi r \alpha a(t-mr))^2(1+(4\pi r \alpha 
a(t-mr))^2)^{-1}d(t)^2. 
\end{equation}
\end{proposition}

@@

  That the reduced objective $\tJa$ is given by \ref{eqn:expjgen} follows from equations \ref{eqn:redexp1}, \ref{eqn:residnormgen}, and
  \ref{eqn:anninormgen}.
\end{proof}

@@

It will be useful to record expressions for the various components of
$\tJa$ when the data is noise-free, that is, the context of
Proposition \ref{thm:norminvexp}, item 3:
\begin{corollary}
  \label{thm:epjnonoise}
  Assume the hypotheses of Proposition \ref{thm:norminvexp}, item
  3. Then noting that $\|d\| = \|w_*\|/(4 \pi r)$
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= \frac{\alpha^4}{2\|w_*\|^2}\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m_*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{(4\pi r)^2}{2\|w_*\|^2}\int \,dt\,  
  \frac{a(t-(m-m_*)r)^2}{(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2)^{2}}w_*(t)^2.
\end{equation}
so 
\begin{equation}
\label{eqn:expjnonoise}
\tJa[m;d] = \frac{(4\pi r \alpha)^2}{2\|w_*\|^2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-1}w_*(t)^2. 
\end{equation}
Finally, if $a \in W^{1,\infty}(\bR)$, then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjnonoise}
  \frac{d}{dm}\tJa[m;d] = -\frac{r (4\pi r \alpha)^2}{\|w_*\|^2} \int \,dt \, 
  \frac{\left(a\frac{da}{dt}\right)(t-(m-m_*)r)}{(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{2}}w_*(t)^2. 
\end{equation}
\end{corollary}

@@

\begin{proposition}
  \label{thm:rampgood}
  Suppose that
  \begin{enumerate}
  \item $m_* \in M$;
  \item $0 < \mu \le \lambda$, and $w_* \in W_{\mu}$;
  \item $d_* = F[m_*]w_*$;
  \item $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
    with $\tau$ given by equation \ref{eqn:taudef}; and
  \item $\alpha > 0$.
  \end{enumerate}
  Then for any $m \in M$, 
  \begin{equation}
    | (m - m_*)r| > \lambda  \Rightarrow  \left|\frac{d}{dm}\tJa[m;d_*]\right| >  
    \frac{r(4 \pi r \alpha)^2(\lambda-\mu)}{(1+(4\pi r\alpha)^2 
      (\lambda+\mu)^2)^{2}}.
    \label{eqn:gradbndnonoise}
  \end{equation}
\end{proposition}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d_*] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore, $a(t) = |t|$, $a a'(t) = t$ in the
  support of the integrand on the RHS of equation
  \ref{eqn:dexpjnonoise}, which therefore 
  becomes (after change of integration variable)
  %%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{d}{dm}\tJa[m;d_*] = -\frac{r (4\pi r\alpha)^2}{\|w_*\|^2} \int \,dt \, 
  t(1+(4\pi r)^2 \alpha^2 
  t^2)^{-2}w_*(t+(m-m_*))^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -\frac{r(4 \pi r \alpha)^2}{\|w_*\|^2}\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, t(1+(4\pi r)^2\alpha^2 t^2)^{-2}w_*\left(t+(m-m_*)r\right)^2.
  \]
  Suppose that $\mu \le \lambda$ and $w_* \in W_{\mu}$. 
  If $m > m_*+\lambda/r$, then $t+(m-m_*)r \in \mbox{supp }w_*$
  implies $-\mu - \lambda < t < \mu-\lambda<0$, so 
  \[
    t(1+(4\pi r)^2\alpha^2 t^2)^{-2} < (\mu-\lambda)(1+(4\pi r)^2\alpha^2 (\mu+\lambda)^2)^{-2}<0
  \]
  in the support of the integrand in equation
  \ref{eqn:gradfinal}. Arguing similarly for $m<m_*-\lambda/r$, obtain
  a similar inequality, implying the conclusion \ref{eqn:gradbndnonoise}.
\end{proof}

@@

The inequality \ref{eqn:estresidnorm} can be interpreted as a bound 
on $\alpha$, given $\epsilon$ and $\lambda$, for a
stationary point of $\tJa$ to yield a solution of the inverse
problem: one obtains a solution, provided that $\alpha$ is
sufficiently small. On the other hand, it is clear that $\alpha$
cannot be too large if stationary points of $\tJa$ are to yield
solutions: the integrand in \ref{eqn:residnorm} is increasing in
$\alpha$ for every $t$ and $m$, and the multiplier
\[
t \mapsto (4\pi r \alpha(t-(m-m_*)r))^4(1+(4\pi r)^2 \alpha^2 
|t-(m-m_*)r|^2)^{-2}
\]
tends monotonically to $1$ as $\alpha \rightarrow \infty$, uniformly
on the complement of any open interval containing
$t=(m-m_*)r$. Therefore
\begin{equation}
  \label{eqn:elimit}
  \lim_{\alpha \rightarrow \infty}e[m,\aw[m;d];d] = \frac{1}{2}.
\end{equation}
Consequently, there exists $\alpha_{\rm max}(\epsilon,\lambda,d)$ so
that
\[
  e[m,\aw[m;d];d]  \le \frac{1}{2}\epsilon^2
  \Rightarrow \alpha \le \alpha_{\rm max}(\epsilon,\lambda,d).
\]
The existence of this limiting penalty weight has been inferred
indirectly. \cite{FuSymes2017discrepancy} describe a constructive
algorithm for its approximation, which is used in the companion paper
\cite{SymesChenMinkoff:21} to dynamically adjust $\alpha$ in the
course of optimization of $\tJa$.
